---
title: "Subsetting Data in R - Lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this lab you can use the interactive console to explore but please record your commands here.  Remember anything you type here can be "sent" to the console with Cmd-Enter (OS-X) or Ctrl-Enter (Windows/Linux) (But only inside the code chunks designated with the ```{r}``` areas).

```{r, message = FALSE}
# don't forget to load the packages that you will need!
library(dplyr)
library(tidyverse)
```

# Part 1

First check to see if you have the `iris` dataset by running this chunk. Remember that code is denoted with backticks "`" around it in the main text, but do not copy backticks when you paste code elements into code chunks (like the gray section below). 

```{r}
iris
```

1. How many observations (rows) are in the `iris` dataset?

```{r}
nrow(iris)
```

2. Take a look at the top and bottom of the data.

```{r}
head(iris)
tail(iris)
```

3. Use `glimpse()` to take a look at the data.

```{r}
glimpse(iris)

```


4. Use `count()` to determine how many of each species of flower are in the data.

```{r}
iris %>% count(Species)

```



5. Select only the Species column

```{r}
iris %>% select(Species)
```

6. Select only the columns that have Species info and Petal info. Create a new data frame called `petal` with this data.

```{r}
petal <- iris %>% select(Species, starts_with("Petal"))
```


7. Pull the data values for Species from the iris dataset.

```{r}
iris %>% pull(Species)
```

4. Filter the rows of `iris` to only those with a Petal.Length greater than 6. What species has these large petals?

```{r}
iris %>% filter(Petal.Length > 6)
```


5. Filter the rows of `iris` to only those that have a `Petal.Length` greater than 6 and a `Petal.Width` greater than 2.

```{r}
iris %>% filter(Petal.Length > 6 & Petal.Width > 2)
```


6. Filter the rows of `iris` to only those that have a `Petal.Length` greater than 6 **OR** a `Petal.Width` greater than 2.

```{r}
iris %>% filter(Petal.Length > 6 | Petal.Width > 2)
```

7. Why do the answers for 5 and 6 differ?

More flowers have either a petal width greater than 2 or a petal length greater than 6 than those that must meet both conditions.


8. Create a new variable called `Petal.Area` that is the `Petal.Length` multiplied by the `Petal.Width`. Hint: use `mutate`.

```{r}
iris %>% mutate(Petal.Area = Petal.Length * Petal.Width)
```
