---
title: "Data Summarization"
output:
  ioslides_presentation:
    css: ../../docs/styles.css
    widescreen: yes
---


```{r, echo = FALSE, message=FALSE, error = FALSE}
library(knitr)
opts_chunk$set(comment = "", message = FALSE)
suppressWarnings({library(dplyr)})
library(readr)
library(tidyverse)
library(jhur)
```

<style type="text/css">
pre { /* Code block - slightly smaller in this lecture */
    font-size: 20px;
}
</style>

# Part 1: Numeric / continuous data

## Data Summarization

* Basic statistical summarization
    * `mean(x)`: takes the mean of x
    * `sd(x)`: takes the standard deviation of x
    * `median(x)`: takes the median of x
    * `quantile(x)`: displays sample quantiles of x. Default is min, IQR, max
    * `range(x)`: displays the range. Same as `c(min(x), max(x))`
    * `sum(x)`: sum of x
    * `max(x)`: maximum value in x
    * `min(x)`: minimum value in x
    
* **all have the ** `na.rm =` **argument for missing data**


## Statistical summarization

The vector getting summarized goes inside the parentheses:

```{r}
x <- c(1, 5, 7, 4, 2, 8)
mean(x)
range(x)
sum(x)
```


## Statistical summarization

Note that many of these functions have additional inputs regarding missing data, typically requiring the `na.rm` argument ("remove NAs"). 

```{r error = TRUE}
x <- c(1, 5, 7, 4, 2, 8, NA)
mean(x)
mean(x, na.rm = TRUE)
quantile(x)
quantile(x, na.rm = TRUE)
```


## Statistical summarization{.codesmall}

You can only do summarization on numeric or logical types. Not characters.

```{r error = TRUE}
x <- c(1, 5, 7, 4, 2, 8)
sum(x)
z <- c("hello", "goodbye")
sum(z)
```


## But how do we do this on dataframes?

First we will need to learn about something called the "pipe".

The pipe is this operator in R: 

`%>%`

It tells R to "pipe" the dataset on the left into the next function.


## Using the pipe `%>%`

```{r}
states <- read_csv("https://hutchdatascience.org/SeattleStatSummer_R/data/states.csv")
states %>% head() # Same as head(states)!
```


## States data

`colnames()` will show us the column names.

```{r}
colnames(states)
```

# Summarizing the data

## Summarize the data: `dplyr` `summarize()` function

`summarize` creates a summary table of a column you're interested in.

<div class = "codeexample">
```{r, eval = FALSE}
# General format - Not the code!
{data to use} %>% 
   summarize({summary column name} = {operator(source column)}) 
```
</div>


## Summarize the data: `dplyr` `summarize()` function

`summarize` creates a summary table of a column you're interested in.

<div class = "codeexample">
```{r, eval = FALSE}
# General format - Not the code!
{data to use} %>% 
   summarize({summary column name} = {operator(source column)}) 
```
</div>

```{r}
states %>% 
  summarize(mean_population = mean(population))
```


## What if there are NAs in my data?

```{r}
states %>% 
  summarize(mean_population = mean(cesarean_percent))
states %>% 
  summarize(mean_population = mean(cesarean_percent, na.rm = TRUE))
```

add `na.rm = TRUE`.


## Summarize the data: `dplyr` `summarize()` function

`summarize()` can do multiple operations at once. Separate by a comma. Breaking line between these keeps things tidy!

```{r}
states %>% 
  summarize(mean_population = mean(population),
            median_population = median(population))
```


## `summary()` Function

Using `summary()` can give you rough snapshots of each numeric column (character columns are skipped):

```{r}
summary(states)
```


# Let's practice!

## Practice

Modify the code below from the `states` dataset to `summarize()` the `fertility_rate_per_1000` column. Find the mean, min, and max.

```{r eval=FALSE}
states %>% 
  summarize(___ = mean(___),
            ___ = min(___),
            ___ = max(___))
```


## Practice

Modify the code below from the `states` dataset to `summarize()` the `fertility_rate_per_1000` column. Find the mean, min, and max.

```{r}
states %>% 
  summarize(mean_fert = mean(fertility_rate_per_1000),
            min_fert = min(fertility_rate_per_1000),
            max_fert = max(fertility_rate_per_1000))
```


## Summary Part 1

- don't forget the `na.rm = TRUE` argument!
- `summary(x)`: quantile information
- `summarize`: creates a summary table of columns of interest

# Part 2: Categorical data

## `count` function

Use `count` to return the number of rows of data.

```{r}
states %>% count()
```


## `count` function

Use `count` to return a frequency table of unique elements of a category (column).

```{r}
states %>% count(state_region)
```


## `count` function

Multiple columns listed further subdivides the count.

```{r, message = FALSE}
states %>% count(state_region, state_division)
```

# Grouping

## Perform Operations By Groups: dplyr

`group_by` allows you group the data set by variables/columns you specify:

```{r}
# Regular data
states
```


## Perform Operations By Groups: dplyr

`group_by` allows you group the data set by variables/columns you specify:

```{r}
states_grouped <- states %>% group_by(state_region)
states_grouped
```


## Summarize the grouped data

It's grouped! Grouping doesn't change the data in any way, but how **functions operate on it**. Now we can summarize `population` by group:

```{r}
states_grouped %>% summarize(total_population = sum(population))
```


## Use the `pipe` to string these together!

Pipe `states` into `group_by`, then pipe that into `summarize`:

```{r}
states %>% 
  group_by(state_region) %>% 
  summarize(total_population = sum(population))
```


# Let's practice!

## Practice

Modify the code to group by `state_region` and summarize by average `fertility_rate_per_1000`.

```{r eval=FALSE}
states %>%
  group_by(___) %>%
  summarize(___ = mean(___))
```


## Practice

Modify the code to group by `state_region` and summarize by average `fertility_rate_per_1000`.

```{r}
states %>%
  group_by(state_region) %>%
  summarize(avg_fert = mean(fertility_rate_per_1000))
```


## Counting

`n()` can also give you the sample size per group (NAs included).

```{r}
states %>% 
  group_by(state_region) %>% 
  summarize(total_population = sum(population),
            sample_size = n())
```

## Summary 

- don't forget the `na.rm = TRUE` argument!
- `summary()`: quantile information
- `summarize`: creates a summary table of columns of interest
- `count(x)`: what unique values do you have?
- `group_by(x)`: changes all subsequent functions
  - combine with `summarize()` to get statistics per group
- `summarize()` with `n()` gives the sample size (NAs included) 

🏠 [Workshop Website](https://hutchdatascience.org/SeattleStatSummer_R/)

