---
title: "Data_Input_Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

In this lab you can use the interactive console to explore or **Knit** the document. 
Remember anything you type here can be "sent" to the console with Cmd-Enter (OS-X) or Ctrl-Enter (Windows/Linux) in an R code chunk.

# Part 1  

1\. Use the manual import method (File > Import Dataset > From Text (`readr`)) to Read in SARS-CoV-2 vaccination data from URL 
http://hutchdatascience.org/intro_to_r/data/vaccinations.csv. You can learn more about how the data was collected here: https://data.cdc.gov/Vaccinations/COVID-19-Vaccinations-in-the-United-States-Jurisdi/unsk-b7fc

You can also run:

```{r}
vacc <- read_csv("https://hutchdatascience.org/SeattleStatSummer_R/data/vaccinations.csv")
```


2\. What is the dataset object called? You can find this information in the Console or the Environment. Enter your answer as a comment using `#`.

```{r}
# vaccinations
```


3\. Preview the data by clicking the table button in the Environment. How many observations and variables are there? Enter your answer as a comment using `#`.

```{r}
# 37272 obs. of 109 variables
```


4\. Use `head()` to display the first few rows of the data frame. 

```{r}
head(vacc)
```


5\. Use `str()` to display the list of column names and their type specification. 

```{r}
str(vacc)
```


6\. Load the `readxl` package with the `library()` command. 

If it is not installed, install it via: `RStudio --> Tools --> Install Packages`. You can also try `install.packages("readxl")`.

```{r}
library(readxl)
```


7\. Download the dataset of asthma prevalence in the USA from: http://jhudatascience.org/intro_to_r/data/asthma.xlsx file to `asthma.xlsx` by running the following code chunk. This only downloads the file, it does NOT bring the file into R.

```{r}
download.file(
  url = "http://jhudatascience.org/intro_to_r/data/asthma.xlsx",
  destfile = "asthma.xlsx",
  overwrite = TRUE,
  mode = "wb"
)
```


8\. Use File > Import Dataset > From Excel to import the data.  What does the code look like? Paste it below.

```{r}
asthma <- read_excel("asthma.xlsx")
```


9\. Use the `head()` and `str()` functions to inspect the data.

```{r}
head(asthma)
str(asthma)
```


10\. Modify the code from #8 to rename the `asthma` object to `my_dataset`.

```{r}
my_dataset <- read_excel("asthma.xlsx")
```
